\myTitle

### 2023-10-21 body-text-of-paper.tex
Matt and Tim's writing

Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given animal \citep{paetkau1995microsatellite,wilson2003bayesian}.  Describing such patterns of movement have many implications, influencing evolutionary processes, disease dynamics \citep{huestis2019windborne}, or patterns of invasive species expansion.

@@

Pairing empirical data with closely derived simulation-based approaches has been long recognized as a tool to help guide the interpretation of empirical data and define the limitations of the available data \citep{vaha2006efficiency,anderson2008improved,latch2011fine}.  Mirroring empirical datasets with forward-in-time simulations often implement sampling with replacement (e.g. citations).  However, sampling with replacement implicitly increases the pairwise genetic distance (i.e., FST) among populations and inflates the perceived ability to correctly assign individuals back to source populations \cite{almudevar2000exact,anderson2008improved}.  \textcolor{red}{Seems like 2 additional sentences here would be appropriate to draw from a study that has demonstrated this process.}  The effects of overestimating our ability to resolve populations creates individuals that discretely align with simulated populations.  These issues could be addressed with simulations that fix observed genetic differences among populations by sampling from the empirical data without replacement (citations for simulation frameworks that use this approach).
  Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be interested in how individuals of admixed ancestry are genetically characterized.  Recognizing that connectivity among genetically distinct populations may be relatively rare, researchers can gain additional insights into the frequency of immigration by also identifying descendants of migrants among study populations (directly [F1 hybrids] or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots, BCX]).  The process of recombination during meiosis creates a distribution of expected associations to parental populations among advanced hybrid classes. Treating linked loci as independent can lead to the false interpretation that the association of advanced hybrid classes to parental population occur strictly as discrete distributions.  Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of linkage among loci available for analysis.
	In response to these needs, we developed gscramble – an approach to simulate genotypes that minimizes bias associated with forward-in-time simulations by sampling alleles from respective populations without replacement, thus preserving the genetic differences observed in the empirical dataset.  Further, by simulating individual genotypes based on user-defined pedigrees, gscramble allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.  By integrating species-specific recombination rates, gscramble simulates biologically informed genotypes that mirror empirical data.  We develop and illustrate gscramble with use of single nucleotide polymorphic (SNP) genotypes.

Can we say something about Anthony Almudevar's visionary method for GSI power assessment \citep{almudevar2000exact}

@@

\subsection*{And a third}

Blah, blah blobbity blah.

>>>>

Genetic clustering algorithms are a commonly used approach to describe genetic structure among populations, reflecting the cumulative effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications, influencing evolutionary processes, maintenance of genetic variation through gene flow, disease dynamics \citep{huestis2019windborne}, or patterns of invasive species expansion CITE.

@@

	A best practice for evaluating the behavior of genetic clustering algorithms on an empirical dataset is to pair the data with closely derived simulations where the truth is known \citep{vaha2006efficiency,anderson2008improved,latch2011fine}. 
	Techniques to simulate individual genotypes frequently use sampling with replacement from the observed allele frequencies.
	However, sampling with replacement implicitly increases the pairwise genetic distance (e.g., FST) among populations and inflates the perceived ability to correctly assign individuals back to source populations \cite{almudevar2000exact,anderson2008improved}.  
	\textcolor{red}{Seems like 2 additional sentences here would be appropriate to draw from a study that has demonstrated this process.}  
	These issues could be addressed with simulations that fix observed genetic differences among populations by sampling from the empirical data without replacement (citations for simulation frameworks that use this approach).

	Beyond simply evaluating the resolution of discrete populations in a simulated framework, researchers may also be interested in how individuals of admixed ancestry are genetically characterized.  
	Recognizing that connectivity among genetically distinct populations may be relatively rare, researchers can gain additional insights into the frequency of immigration by also identifying descendants of migrants among study populations (directly [F1 hybrids] or advanced hybrids reflecting the effect of backcrossing across multiple generations [BC1, \ldots, BCX]).  
	The process of recombination during meiosis creates a distribution of expected associations to parental populations among advanced hybrid classes. 
	Treating linked loci as independent can lead to the false interpretation that the association of advanced hybrid classes to parental population occur strictly as discrete distributions.  
	Therefore, in evaluating descendants of migrants within a simulated framework, it is imperative to take into consideration the pattern of linkage among loci available for analysis.

	In response to these needs, we developed gscramble – a simulation approach that samples alleles from respective populations without replacement, thus maintaining the genetic differences in the empirical dataset.  
	Further, by simulating individual genotypes based on user-defined pedigrees, gscramble allows for the simulation of admixed genotypes of varying degrees of complexity while allowing the tracking of haplotype blocks from source populations within the specified pedigree.  
	By integrating species-specific recombination rates, gscramble simulates biologically informed genotypes that mirror empirical data.  
	We develop and illustrate gscramble with use of single nucleotide polymorphic (SNP) genotypes.

@@

\subsection*{Empirical Example}

Genetic samples were collected from invasive wild pigs across a SPATIAL EXTENT in southeastern Missouri that were lethally removed through population control efforts conducted as a component of the Missouri Feral Hog Elimination Partnership 
by US Department of Agriculture – Animal Plant Health Inspection Services – Wildlife Services, Missouri Department of Conservation, and other cooperative state agencies CHECK. 
DNA was extracted from hair or tissue (collected from animals at the time of euthanasia) with commercially available magnetic bead recovery kits (MagMax DNA, Thermo Fisher Scientific).  
Genotyping was performed with GeneSeek’s Genomic Profiler for Porcine HD which provides 62,128 biallelic SNP loci that are mapped across the 18 autosomal chromosomes. 
We implemented standard genotype quality control metrics, specifically, we pruned genotypes with call rates <95\% and minor allele frequencies <5\%.
We implemented linkage disequilibrium (LD) pruning using a window size of 50 loci and step size of 5 loci to remove markers above a linkage threshold of R2 > 0.5.

GSCRAMBLE REFERENCE GROUPS
GSCRAMBLE PEDIGREES
ADMIXTURE GSCRAMBLE VS EMPIRICAL
FST OF EMPIRICAL AND GSCRAMBLE CORE POPULATIONS
